---
title: "Part2: Thought Probe Analysis"
output: html_document
date: "2023-04-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### A bit of exploratory analysis of the thought probe data from Kassandra's first 20 participants.
##### Goal: to judge if participants' performance errors occur because of off-task episodes and if off-task episodes correspond to performance errors. 

```{r}
data <- read.csv('first_participants_probes.csv')
data <- data[,-1]
head(data)
```


## Probe 1: Participants Self-Rating of Ontask Level (Scale of 1-6)

### 1. What does the distribution of participant ratings look like? 

```{r}
#Read in data 
data <- read.csv('first_participants_probes.csv')
library(ggplot2)
data$probe1 <- as.integer(data$probe1)
ggplot(data=data, aes(x=probe1)) + geom_bar() + xlab("Participants' Ontask Rating")
```
Overall, participants tend to report high levels of on-task-ed-ness (which makes sense, participants usually try to do their best on the task). So now, let's see how often they remember the correct sequence of letters based on their on-task rating. 

```{r}
data$trial_correct <- ifelse(data$trial_score == 1.0, TRUE, FALSE)
ggplot(data=data, aes(x=probe1, fill=trial_correct)) + geom_bar() + xlab("Participants' Ontask Rating")
```
It seems that, a small fraction of the time, participants make mistakes even when they are fully focused. Conversely, sometimes participants complete the trial correctly even when they admit to being somewhat distracted during the trial. This is especially notable for ratings of "4" and seems to hint that participants are able to be a bit unfocused and still complete the trial without (letter) errors. 

But of course, the above bar plot only shows a binary measure. To explore trial performance as a continuous measure, see the plots below. 

```{r}
data$probe1 <- as.character(data$probe1)
ggplot(data, aes(y=trial_score, x=probe1)) + geom_boxplot() + geom_violin() + xlab("Participants' Ontask Rating") + ylab("Trial Percentage Correct")
```

```{r}
data$probe1 <- as.integer(data$probe1)
ggplot(data, aes(y=trial_score, x=probe1, color = as.character(time_condition))) + geom_point() + geom_jitter() + xlab("Participants' Ontask Rating") + ylab("Trial Percentage Correct") + labs(color = "Time Condition (in seconds)")

```

```{r}
linreg <- lm(probe1~trial_correct, data=data)
summary(linreg)
```
```{r}
ggplot(data, aes(y=trial_score, x=probe1)) + geom_point() + geom_jitter() + xlab("Participants' Ontask Rating") + ylab("Trial Percentage Correct") +  stat_smooth(method="lm")
```

There's clearly not a strong linear relationship (note the very small R squared value)... perhaps a psychometric curve would fit the data better? 


### 2. How does trial duration relate to ontask rating? 

Do participants report lower on-task ratings during the longer (30 second) time condition? 

```{r}
ggplot(data=data, aes(x=probe1, fill=as.character(time_condition))) + geom_bar(position="dodge") + xlab("Participants' Ontask Rating") + labs(fill = "Time Condition (in seconds)")
```
There doesn't seem to be much difference in ratings by time condition. 

The "4" on-task rating is especially intersting to me because there's so much variability in performance. Perhaps that's due to time condition: maybe participants are more likely to complete the trial successfully in longer trials even if they get distracted because there's more opportunities for rehearsal?


```{r}
correct_four <- data$probe1 == 4 & data$trial_score == 1
length(data$time_condition[correct_four == TRUE & data$time_condition == 20]) #number of correct trials for 20-second durations in which participants selected "4" on the thought probe scale
length(data$time_condition[correct_four == TRUE & data$time_condition == 30]) #number of correct trials for 30-second durations in which participants selected "4" on the thought probe scale
```

```{r}
incorrect_four <- data$probe1 == 4 & data$trial_score != 1
length(data$time_condition[incorrect_four == TRUE & data$time_condition == 20]) #number of incorrect trials for 20-second durations in which participants selected "4" on the thought probe scale
length(data$time_condition[incorrect_four == TRUE & data$time_condition == 30]) #number of incorrect trials for 30-second durations in which participants selected "4" on the thought probe scale
```

Time condition truly doesn't seem to make a difference. 

### Conclusions: participants are sometimes are able to do the task correctly even when they are a little distracted. But this may vary from participant to participant. Further exploration of participants' individual data is needed. 




